---
title: "House_Stark_Network"
author: "ADAMU Farida"
date: "2023-04-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Analysing the House Stark main characters in the Game of Thrones Series 
Game of Thrones is a popular TV show  where Nine noble families wage war against each other in order to gain control over the mythical land of Westeros. In this analysis, we focus on the major characters in House Stark - Arya, Jon and Sansa Stark. 


```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide'}
library(igraph)
library(dplyr)
```


```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide'}
### Read the data from external files
got_nodes <- read.csv(file="got-nodes.csv")
got_edges <- read.csv(file="got-edges.csv")
```


```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide'}
#subset the datarframe for only the main characters in House Stark 
#We select only characters that made it to the very final episodes. 
edges <- got_edges %>% filter(Source == "Arya" | 
                                Source == "Jon" | 
                                Source == "Sansa")
```


```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide'}
### Generate a network from the dataframes
network <- graph.data.frame(edges,
                           # vertices = got_nodes,
                            directed=TRUE)
is.igraph(network)
```

```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide'}
#information about the vertices
V(network)

#information about the edges or connections
E(network)

#labels
V(network)$label <- V(network)$name
V(network)$label

#number of degrees
V(network)$degree<- degree(network)
V(network)$degree

```


First, we plot a histogram of the degree of the nodes in the network of these three characters. The degree of a node is the measure of how many neighbors it has in the network. 

```{r echo=FALSE, warning=FALSE, message=FALSE}
#histogram
hist(V(network)$degree,
     col = 'grey',
     main = 'histogram of node dgree',
     ylab = 'frequency',
     xlab = 'degree of vertices')
#the histagram shows us thre are more characters with smaller degrees
#than with higher degrees in the game of thrones series.
#there are few nodes with high connections and many notes with small connecyions
```

From the histogram  we see that there are few nodes with high connections(degrees) and many nodes with small connections. We will explore which of the characters has the most connections shortly, in the network analysis. 

```{r echo=FALSE, warning=FALSE, message=FALSE}
# Plot network with vertex labels
plot(network, layout = layout.fruchterman.reingold,
     vertex.size = V(network)$degree, 
     vertex.label.dist= 0.5,
     edge.arrow.size = 0.3,
    vertex.label.cex = 0.5)
```


Robert is the most central node or point of connection between Arya, Jon and Sansa and this is very simple to explain because he is their father. However we see he has a very small node as he dies very early in the series. We also see that each of these characters despite being siblings their own interaction circles as they get seperated from one another after their dad was publicly executed.  Arya and Sansa have some characters they bothe interacted with at some point in the series, Cersei who was supposed to Sansa's mother-in-law and Bran who is their brother but also died early on in the series. 



## Detecting Communities

The Louvain Algorithm

There’s a unique connection between Tyrion and Sansa even though they are both in different communities. This is because somewhere in the series, they begin to interact after Tyrion saves Sansa from a forced marriage. Sansa and Arya also have connections with Jon, but Jon seems to be isolated from both of them as the only connections between him and his sisters are with Robert. 
```{r echo=FALSE, warning=FALSE, message=FALSE}
## Communities according to the Louvain algorithm
louvain <- cluster_louvain(as.undirected(network))
plot(network, layout=layout.fruchterman.reingold,
     vertex.size=4, vertex.label.dist=0.5,
     vertex.color=membership(louvain),
     edge.arrow.size=0.5)
```

### The Edge- Between Algorithm

Sansa and Arya have more connections between them than Jon has with any of them. 

```{r, echo=FALSE, warning=FALSE, message=FALSE}

##C Communities accroding to the edge-between. algorithm
edge_between<-cluster_edge_betweenness(as.undirected(network))
plot(network, layout=layout.fruchterman.reingold, vertex.size=4, 
     vertex.label.dist=0.5, vertex.color=membership(edge_between), edge.arrow.size=0.5)
```

### Densely connected nodes

Sansa has the strongest connections in the network and Jon has the weakest. In the end, this pays off for her as she becomes queen in the north. 


```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Community detection to detect densely connected nodes
unet <- graph.data.frame(edges, directed = F)
cnet <- cluster_edge_betweenness(network)
plot(cnet,
     unet,
     vertex.size = V(network)$degree,
     vertex.label.cex = 0.5,
     vertex.label.dist=0.5, 
    edge.arrow.size=0.3)
```

```{r echo=FALSE, warning=FALSE, message=FALSE}
# Create a data frame with network statistics
network_stats <- data.frame(
  Nodes = vcount(network),
  Edges = ecount(network),
  MeanDegree = mean(degree(network, mode ="all")),
  SDDegree = sd(degree(network, mode ="all")),
  Isolates = sum(degree(network, mode ="all")==0),
  AvgClustering = transitivity(network, type="global")
)

# View the network statistics table
library(knitr)
# Create a table with the network statistics
kable(network_stats)

```


```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide'}
## Distances 
diameter(network); mean_distance(network, directed=F)

### Connectivity
indegree<-degree(network, mode="in")
outdegree<-degree(network, mode="out")
degree <-degree(network, mode="all")

reciprocity(network)

```

Distance; it takes about 2.45 (2.447843) steps to get from one node to another in the network graph

Reciprocity; We have a reciprocity of 0 which means there are no mutual edges in the graph.

```{r echo=FALSE, warning=FALSE, message=FALSE, results ='hide'}
## Assortativity
## ...using simple correlation coefficient
es <-get.edges(network, E(network))
deg<-degree(network, mode="all")
assortativity= cor(deg[es[,1] ], deg[es[,2] ])
```

Assortativity;  We have an assortativity of  -0.08346024. This indicates that there is a slight tendency for nodes in the graph to connect to nodes with different attributes rather than nodes with similar attributes.


```{r echo=FALSE, warning=FALSE, message=FALSE, results ='hide'}
## Positioning:
# Global centrality
betweenness= betweenness(network, directed = FALSE)
betweenness

eigenvector.centrality= evcent(network)
hist(eigenvector.centrality$vector)
```

Sansa has the highest score of 1.0,  making her the most important node in this network. Arya follows closely with 0.8404530 , then Jon, then others. 

```{r echo=FALSE, warning=FALSE, message=FALSE, results ='hide'}
# Local centrality
outdegree= degree(network, mode="out")
indegree= degree(network, mode="in")
deg= degree(as.undirected(network))
```


```{r echo=FALSE, warning=FALSE, message=FALSE, results ='hide'}
# Clustering coefficient
clustering = transitivity(as.undirected(network), type="local")
clustering
```
Clustering; Clustering is only detected between the three main characters, other characters they interact with outwardly never get to meet the other characters in-between them. 

## References

Data Source - https://github.com/melaniewalsh/sample-social-network-datasets/tree/master/sample-datasets/game-of-thrones

Netwrok of Thrones Idea - https://www.maa.org/sites/default/files/pdf/Mathhorizons/NetworkofThrones%20%281%29.pdf

